{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2024-03-18 paper/main.tex
add weak selection comments

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\! 1\). 
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

@@

We note that the expected \emph{total} payoff in the game, $u_{0}+u_{1}+\cdots +u_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}u_{t}$. This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}u_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$). Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature. On that point, we note that many of the differences between realized and expected payoffs disappear in the limit of weak selection \citep{mcavoy:NHB:2020}. Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.

>>>>

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\! 1\)). This is not a coincidence, since it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

@@

We note that the expected \emph{total} payoff in the game, $u_{0}+u_{1}+\cdots +u_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}u_{t}$. This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}u_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$). Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature. On that point, we note that differences between realized and expected payoffs disappear in the limit of weak selection, which is known in a general setting \citep{mcavoy:NHB:2020}. Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.
